\textls[-0]{\thetitle}

### 2023-03-04 casestudies.tex
Change according to Reviewer Please Review!!!

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to the metatheories, this case study produces abstract
interpreters that are directly ready for program extraction.

@@

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
ready for program extraction to OCaml.
We put them to test over queries.

@@

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for them.
%\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

>>>>

Our second case study is a mechanization of abstract interpreters
for simple imperative languages.
In addition to the proof of its soundness, this case study produces abstract
interpreters that are directly ready for program extraction.

@@

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
ready for program extraction to OCaml.
We put the verified abstract interpreters to test over several simple queries. % Reviewer D : l899

@@

In addition to the two case studies above, we also use extensible
inductive types for modeling extensible context-free grammars and derive
decision procedures for language membership checking.
%\EDJ{I will suggest change to ``and derive decision procedures for them (based on naive parsing by enumeration)''.}

### 2023-03-04 challenge.tex
Change according to Reviewer Please Review!!!

\noindentparagraph{%
\headinglabel{C1}{chg:extensible-v-exhaustive}.\,
Extensible inductive types vs.\ exhaustive induction reasoning.%
}

@@

However, there is a tension between extensibility of inductive types
and exhaustivity of induction reasoning.\footnote{%
The tension reflects a duality between variants and records.
With record-like language constructs (e.g., OO classes),
it is safe for an extension to add new fields:
existing fields can still be projected.
But variant-like constructs (e.g., inductive types) do not automatically
enjoy safe, modular addition of constructors:
existing pattern matches could become non-exhaustive.}
In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
the substitution function, require induction (i.e., elimination of inductive types).
A language design must enforce that induction remains exhaustive in the
face of the new constructors in the derived family.
For modularity, the type system should do so without requiring
redefinition or rechecking of those cases already handled by the base
family.

@@

Family polymorphism enables modular reuse via late binding:
code of a base family can be reused by a derived family,
because fields referenced by that code have meanings polymorphic to the
enclosing family.

This flexibility, however, prevents \emph{definitional equality}
that one takes for granted when programming in a proof assistant.
In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
polymorphism cannot equate references to the substitution function and
its definition (i.e., a pattern match against four cases),
as a derived family may modify the definition of the function by adding
new cases.
Without the ability to unfold the substitution function, how can the
programmer even prove the substitution lemma?
%
The problem is compounded by the occasional need in derived families to override fields,
which is potentially at odds with being able to use equalities over the fields.

>>>>

\noindentparagraph{%
\headinglabel{C1}{chg:extensible-v-exhaustive}.\,
Extensible inductive types vs.\ exhaustive inductive reasoning.% Reviewer D
}

@@

However, there is a tension between extensibility of inductive types
and exhaustivity of inductive reasoning.\footnote{%
The tension reflects a duality between variants and records.
With record-like language constructs (e.g., OO classes),
it is safe for an extension to add new fields:
existing fields can still be projected.
But variant-like constructs (e.g., inductive types) do not automatically
enjoy safe, modular addition of constructors:
existing pattern matches could become non-exhaustive.}
In \cref{fig:stlc-nonmechanized}, all the lemmas and theorems, as well as
the substitution function, require induction (i.e., elimination of inductive types).
A language design must enforce that induction remains exhaustive in the
face of the new constructors in the derived family.
For modularity, the type system should do so without requiring
redefinition or rechecking of those cases already handled by the base
family.

@@

Family polymorphism enables modular reuse via late binding:
the code of a base family can be reused by a derived family,
because fields referenced by that code have meanings polymorphic to the
enclosing family.

This flexibility, however, prevents \emph{definitional equality}
that one takes for granted when programming in a proof assistant.
In \cref{fig:stlc-nonmechanized}, a proof assistant supporting family
polymorphism cannot unfold references to the substitution function into its definition (i.e., a pattern match against four cases), %Review D : l135
as a derived family may modify the definition of the function by adding
new cases.
Without the ability to unfold the substitution function, how can the
programmer even prove the substitution lemma?
%
The problem is compounded by the occasional need in derived families to override fields,
which is potentially at odds with being able to use equalities over the fields.

### 2023-03-04 coqimpl.tex
Change according to Reviewer Please Review!!!

Consider the translation of \lsti{subst} in family \lsti{STLC}.
First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
every time the programmer completes a \lsti{Case}, a case handler (e.g., \lsti{Def subst\_tm_unit})
is generated and added to the module.

>>>>

Consider the translation of \lsti{subst} in family \lsti{STLC}.
First, a module named \lsti{STLC\_subst\_Cases} is generated interactively:
every time the programmer completes a \lsti{Case}, the case handler (e.g., \lsti{Def subst\_tm_unit})
is type-checked and added to the module. % Reviewer D: l487

### 2023-03-04 introduction.tex
Change according to Reviewer Please Review!!!

The EP is well studied in the conventional setting of functional
programming and object-oriented (OO) programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP.

>>>>

The EP is well studied in the conventional setting of functional
programming and object-oriented (OO) programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP.\EDJ{Reviewer D wants citation here}

### 2023-03-04 metatheory-condensed.tex
Change according to Reviewer Please Review!!!

We use Tarski-style universes~\cite{hofmann1997syntax}: a universe~$\cU$
is inhabited by the \emph{codes} of types,
with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
and $\codety{T}$ encoding type~$T$ (rule \ruleref{tm/c}).
MLTT has an infinite hierarchy of universes;
we omit universe levels in the presentation for conciseness.

@@

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
A constructor is given by a pair of types
$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
$\Wsup{\wsig}{t_1}{t_2}$ constructs a term of type~$T$ (rule \ruleref{tm/wsup}).
As the rules suggest, our formulation additionally supports
\emph{W-type signatures}.
A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
composed of $n$ pairs of types, each of which represents a constructor
(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).

>>>>

In dependent type system, types are treated as terms, and should have a type. % Reviewer A : Why is W(t) a term and not a type?
We use Tarski-style universes~\cite{hofmann1997syntax}: every type has a \emph{code}, inhabits in a universe~$\cU$,
with $\El{t}$ decoding~$t$ (rule \ruleref{ty/el})
and $\codety{T}$ encoding type~$T$ (rule \ruleref{tm/c}).
MLTT has an infinite hierarchy of universes;
we omit universe levels in the presentation for conciseness.

@@

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
A constructor is given by a pair of types
$\goodType{\Gm}{A}{}$ and $\goodType{\Gm,A}{B}{}$ (rule \ruleref{wsig/add}):
given two arguments $\goodTerm{\Gm}{t_1}{A}$ and
$\goodTerm{\Gm,\sub{B}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{T}$,
we can constructs a term $\Wsup{\wsig}{t_1}{t_2}$ of type~$T$ (rule \ruleref{tm/wsup}). % Reviewer B: Line 666
As the rules suggest, our formulation additionally supports
\emph{W-type signatures}.
A well formed W-type signature \goodWSig{\Gm}{\wsig}{n} is
composed of $n$ pairs of types, each of which represents a constructor
(rules \ruleref{wsig/empty} and \ruleref{wsig/add}),
with $\wsigproj{i}{1}{\wsig}$ and $\wsigproj{i}{2}{\wsig}$ giving the
$i$-th pair in~$\wsig$, and $\wcode{\wsig}$ giving the code of the W-type (\ruleref{tm/w}).

### 2023-03-04 related-work.tex
Change according to Reviewer Please Review!!!

\citet{delaware2011} engineer product lines of theorems and proofs built
from feature modules. Feature composition is manual, which seems to have
motivated later approaches based on DTC.

The design pattern \emph{data types à la carte} (DTC) introduced for Haskell by~\citet{dtc}
is a source of inspiration for many approaches
\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
%\cite{vdr2022intrinsically}
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.

>>>>

\citet{delaware2011} engineer product lines of theorems and proofs built
from feature modules. Feature composition is manual, which seems to have
motivated later approaches based on \emph{data types à la carte} (DTC)~\cite{dtc}. % Reviewer B : Line 910
DTC has in turn served as a source of inspiration for numerous work
\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
%\cite{vdr2022intrinsically}
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.
